---
title: "Group_11_Analysis"
format: html
editor: visual
---

# EDA Part

## Introduction

Coffee quality is a key link connecting producer profits and consumer experience. It not only directly determines market value, but also affects the sustainable development of the coffee industry. This study relies on the real production records of the Coffee Quality Database (CQD), focusing on the three core sensory indicators of aroma, flavor, and acidity, as well as key production characteristics such as the number of category_tow_defects, average planting altitude, and harvest year. We conducted a systematic exploratory analysis of the data and constructed a generalized linear model (GLM), which clearly revealed the direction and strength of the impact of each feature on coffee quality classification. The research results provide guidance for producers to help improve coffee quality.

```{r}
#| echo: false
#| warning: false
#| message: false
#load the packages
library(tidyverse)
library(naniar) 
library(ggplot2)
library(tidyr)
library(skimr)
library(knitr)
library(dplyr)
library(kableExtra)
library(gridExtra)
library(grid)
library(GGally)
```

```{r data cleaning}
#| echo: false
#| warning: false
#| message: false
#| results: 'hide'
#| include: false
data <- read.csv("C:\\Users\\29583\\Downloads\\dataset11.csv")
glimpse(data)
summary(data)
#check the information of missing values
miss_var_summary(data)

#remove missing values
coffeedata <- data %>%
  drop_na(harvested, country_of_origin,altitude_mean_meters)
miss_var_summary(coffeedata)

#find the unreasonable values in altitude_mean_meters
ggplot(coffeedata, aes(x = 1, y = altitude_mean_meters)) +
  geom_point(size = 1.5)

#It is clear that there are two points higher than 100000 meters, which is unreasonable for planting coffee. So we have to remove them.
coffeedata <- coffeedata%>%
  filter(altitude_mean_meters < 10000 | is.na(altitude_mean_meters))

#histogram and box plot about the distribution and outliers of altitude_mean_meters
ggplot(coffeedata, aes(x = altitude_mean_meters)) +
  geom_histogram(bins = 30, color = "black", fill = "gray80") +
  labs(title = "Distribution of Coffee Planting Altitude",
       x = "Altitude (meters)",
       y = "Frequency") +
  theme_minimal()
#from the distribution of altitude_mean_meters, on the right side (approximately greater than 2500 meters), there is a long tail feature and the sample size is extremely small.
ggplot(coffeedata, aes(y = altitude_mean_meters)) +
  geom_boxplot(fill = "gray80", width = 0.3) +
  geom_hline(yintercept = 2500, color = "red", linetype = "dashed", linewidth = 1) +
  labs(title = "Boxplot of Coffee Planting Altitude",
       y = "Altitude (meters)") +
  theme_minimal() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
sum(coffeedata$altitude_mean_meters >= 2500, na.rm = TRUE)

#Here we found through research that the maximum height for conventional coffee cultivation is 2500 meters. From the box plot we can also find there are only 12 samples over 2500 meter, which are outliers, so we can delete them.
coffeedata <- coffeedata %>%
  filter(altitude_mean_meters < 2500 | is.na(altitude_mean_meters))

#Considering the impact of variable numerical scale on model estimation, the altitude is converted from meters to kilometers to reduce the difference in variable magnitude.
coffeedata$altitude_km <- coffeedata$altitude_mean_meters / 1000

#remove unnecessary columns
coffeedata <- coffeedata %>% 
  select(-country_of_origin, -altitude_mean_meters)

#For logistic regression models, the dependent variable should be of factor type. but the Qualityclass is chr type, so we have to turn it into factor.
coffeedata$Qualityclass <- factor(
  coffeedata$Qualityclass,
  levels = c("Poor", "Good")
)
str(coffeedata)

```

We systematically processed the original coffee dataset to ensure the reliability of subsequent modeling and analysis.

Firstly, we check the data structure and missing values, and remove invalid samples with missing key variables. Then, outlier screening was performed on the average altitude. Through visualization, it was found that there were two extreme values exceeding 100000 meters, which obviously do not match the actual coffee cultivation and have been removed. Further examination through histograms and boxplots revealed that there were 12 extreme samples with elevations exceeding 2500 meters, which were also deleted.

In order to standardize the numerical scale, we converted the average altitude from meters to kilometers. Subsequently, unnecessary columns were removed and the coffee quality variable was converted into a factor variable.

## Exploratory Analysis

```{r description table}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-description-of-Qualityclass
#| tbl-cap: Description of Qualityclass
#| 
#description of Qualityclass
coffeedata %>%
  count(Qualityclass, name = "Frequency") %>%  
  mutate(Variable = "Qualityclass") %>%       
  select(Variable,Category = Qualityclass, Frequency) %>%
  kable(align = "ccc", booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position", font_size = 11) %>%
  row_spec(0, align = "center")
```

@tbl-description-of-Qualityclass shows the frequency of categorical variable. The variable Qualityclass is a binary categorical variable, consisting of two levels: "Good" (high-quality coffee) and "Poor" (low-quality coffee). The sample distribution is highly balanced, with 444 high-quality coffee samples (accounting for 50.5%) and 436 low-quality coffee samples (accounting for 49.5%). There is no problem of category imbalance, which can ensure the stability of subsequent models.

```{r summary statistics}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-summary-of-quantitative-variables
#| tbl-cap: summary of quantitative variables
#summary of the quantitative variables
vars <- c("aroma", "flavor", "acidity", "category_two_defects","harvested","altitude_km")
data.frame(
  Mean = apply(coffeedata[vars], 2, function(x) round(mean(x, na.rm=T), 2)),
  Median = apply(coffeedata[vars], 2, function(x) round(median(x, na.rm=T), 2)),
  SD = apply(coffeedata[vars], 2, function(x) round(sd(x, na.rm=T), 2)),
  Min = apply(coffeedata[vars], 2, function(x) round(min(x, na.rm=T), 2)),
  Max = apply(coffeedata[vars], 2, function(x) round(max(x, na.rm=T), 2)),
  IQR = apply(coffeedata[vars], 2, function(x) round(IQR(x, na.rm=T), 2))) %>%
  rownames_to_column("Variable") %>%
  kable(align = "ccccccc",booktabs=TRUE) %>%
  kable_styling()
```

@tbl-summary-of-quantitative-variables shows the descriptive statistics of the key quantitative variables of coffee samples, including three sensory quality indicators, one defect count indicator, one harvest year indicator and one indicator of average planting altitude.

Aroma scores fall in a moderate-to-high range with a mean value of 7.57 and a median value of 7.58, indicating a highly similar distribution. The standard deviation is only 0.32, with minimal fluctuations, indicating that there is not much difference in aroma scores between samples, and the overall concentration is at a relatively high level.

Flavor scores is Similar to the aroma performance, which remains stable and concentrated at a medium to high level.

Acidity also maintains a moderate-to-high scoring level, with a mean of 7.54 and a median of 7.58, while the standard deviation is 0.31, which is the smallest fluctuation among the three sensory indicators, indicating that acidity is the most stable performance.

The mean number of category_tow_defects is 3.51, but the median is only 2.00, the standard deviation is as high as 5.24 and the maximum value reaches 45.00, indicating a significant right bias in the data. Most samples have fewer defects, but there are a few extreme samples with extremely many defects.

The harvest years are concentrated between 2010 and 2018, with a standard deviation of 1.81, indicating a small sample time span and compact year distribution.

The average altitude is 1.30km, the median is 1.31km, and the distribution is approximately symmetrical. The standard deviation is 0.41, with a maximum value of 2.14km and a minimum value of 0.00km, indicating that the sample covers coffee planting areas from low to high altitudes, and is generally concentrated at a suitable planting height of around 1.3km.

In summary, the mean and median of the three sensory indicators are close, the standard deviation and IQR are small, and the distribution is symmetrical and concentrated. The categorization shows a right skewed feature.

The above is a descriptive statistical summary of the research data. Based on this, visual analysis is used to further visualize the distribution patterns and data characteristics of each variable.

```{r pairplot}
#| label: fig-pairplot
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-cap: pair plot of variables
#a pair plot to show the correlation and distribution of these variables. At first it contains the variable 'harvested', but it shows non-relationship to other variables and in order to keep the ensure the cleanliness of the image, we remove 'harvested' in the pair plot
coffeedata %>%
  select(aroma, flavor, acidity,altitude_km,category_two_defects) %>%
  ggpairs()+
  theme_bw()+
  theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_blank())
```

This pair plot comprehensively presents the distribution and correlations of these variables.

The diagonal density curve shows that the three sensory indicators of aroma, flavor, and acidity exhibit an approximately symmetrical unimodal distribution. There is a double peak in the harvest year. The number of category_two_defects exhibits a clear right skewed long tail feature.

The correlation coefficients between aroma, flavor, and acidity are all above 0.6, indicating a strong positive correlation among the three. The average altitude is weakly positively correlated with sensory indicators, while the defects is weakly negatively correlated with sensory indicators Overall, there may be collinearity issues within sensory indicators.

```{r boxplot}
#| label: fig-boxplot-sensory-quality-indicators
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-cap: Boxolot of aroma,flavor and acidity
# Function of getting legend from a plot
get_legend <- function(p){
  g <- ggplotGrob(p)
  for (i in seq_along(g$grobs)){
    if (g$grobs[[i]]$name == "guide-box"){
      return(g$grobs[[i]])
    }
  }
}
#create a color for the plot
my_color <- c("#4292db", "#E8785A")
aroma_box <- coffeedata %>%
  ggplot(aes(x = factor(Qualityclass,
                        labels=c("Good","Poor")),
             y = aroma,
             fill=factor(Qualityclass)))+
  scale_fill_manual(values = my_color)+
  geom_boxplot(alpha=0.9)+
  labs(x="Qualityclass", y="Aroma", fill="Qualityclass")+
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray90"),
        axis.title = element_text(size = 8))

flavor_box <- coffeedata %>%
  ggplot(aes(x = factor(Qualityclass,
                        labels=c("Good","Poor")),
             y =flavor,
             fill=factor(Qualityclass)))+
  scale_fill_manual(values = my_color)+
  geom_boxplot(alpha=0.9)+
  labs(x="Qualityclass", y="Flavor", fill="Qualityclass")+
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray90"),
        axis.title = element_text(size = 8)) 

acidity_box <- coffeedata %>%
  ggplot(aes(x = factor(Qualityclass,
                        labels=c("Good","Poor")),
             y = acidity,
             fill=factor(Qualityclass)))+
  scale_fill_manual(values = my_color)+
  geom_boxplot(alpha=0.9)+
  labs(x="Qualityclass", y="Acidity", fill="Qualityclass")+
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray90"),
        axis.title = element_text(size = 8))

# Get a legend
legend_boxplot <- get_legend(aroma_box)

# Remove all the legends
aroma_box <- aroma_box + theme(legend.position = "none")
flavor_box <- flavor_box + theme(legend.position = "none")
acidity_box <- acidity_box + theme(legend.position = "none")

# Create a panel
panel_coffee_boxplot <- arrangeGrob(aroma_box, flavor_box, acidity_box, ncol = 3)

# Draw the plot
grid.draw(panel_coffee_boxplot)
```

This set of boxplots clearly shows the distribution differences between high-quality and low-quality coffee in three sensory indicators. The aroma, flavor, and acidity scores of high-quality coffee are generally higher, and the data is more concentrated and stable. Low quality coffee has lower scores, greater fluctuations, and more outliers in low scores. The characteristics of the two sets of outliers are also significantly different, with higher values in the high-quality group and lower values in the low-quality group. Overall, these sensory indicators can effectively distinguish coffee quality and provide intuitive basis for subsequent statistical analysis.

```{r histogram of average planting altitude}
#| label: fig-histogram-altitude
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-cap: histogram of altitude_km
#draw a histogram of average planting altitude to show the distribution 
ggplot(coffeedata, aes(x = altitude_km, fill = Qualityclass)) +
  geom_histogram(alpha = 0.9, bins = 20,
                 position = "dodge",
                 color = "white", linewidth = 0.1) +
  scale_fill_manual(values = my_color) +
  labs(x = "Altitude (kilometers)",
       y = "Frequency",
       fill = "Quality") +
  theme_minimal(base_size = 14)
```

@fig-histogram-altitude performs that altitude distribution of coffee of different qualities shows significant differences. High quality samples are mainly concentrated at altitudes of 1.5-1.8 kilometers, while low-quality samples are more concentrated at altitudes of 1.0-1.3 kilometers, with an overall bias towards low altitudes. The proportion of high-quality coffee is higher in high-altitude (\>1.5 kilometers) areas, while the frequency of low-quality coffee is higher in low altitude (\<1.2 kilometers) areas. This indicates that the higher the altitude, the more likely the coffee quality is to improve, providing intuitive evidence for the positive correlation between altitude and coffee quality.

```{r histogram of defects}
#| label: fig-histogram-defects
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-cap: histogram of defects
#draw a histogram of category_two_defects to show the difference distribution of the defects number in different coffee qualities.
ggplot(coffeedata, aes(x = category_two_defects, fill = Qualityclass)) +
  geom_histogram(alpha = 0.9, bins = 20,
                 position = "dodge",
                 color = "white", linewidth = 0.1) +
  scale_fill_manual(values = my_color) +
  labs(x = "defects of two category",
       y = "Frequency",
       fill = "Quality") +
  theme_minimal(base_size = 14)
```

@fig-histogram-defects shows the distribution of the number of category_two_defects in coffee of different qualities varies significantly. Overall, the number of defects shows a right skewed distribution, with most samples concentrated in the low range of 0-5 defects, and a peak formed on the left side. High quality coffee is almost entirely concentrated on 0-5 defects, and as the number of defects increases, the sample frequency rapidly decreases. In contrast, low-quality coffee has a wider distribution of defects, with samples containing 0-10 defects, and a significantly higher number of high defect segments. This difference clearly indicates that the number of defects can effectively distinguish coffee quality, and samples with fewer defects are more likely to belong to the high-quality group.

```{r}
#| label: fig-bar-harvested
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-cap: harvested stacked percentage bar plot
#draw the stacked percentage bar plot to show the proportion of good coffee and poor coffee in each year and explore the relationship between harvested year and coffee quality
ggplot(coffeedata, aes(x = as.factor(harvested), fill = Qualityclass)) +
  geom_bar(position = "fill", alpha = 0.9,width = 0.6) +
  scale_fill_manual(values = my_color) +
  labs(x = "Harvest Year",
       y = "Proportion",
       fill = "Quality") +
  theme_minimal()
```

@fig-bar-harvested shows the proportion distribution of coffee quality in each year from 2010 to 2018. From the plot, it can be seen that the quality of coffee fluctuates greatly in different years and lacks a stable trend. For example, in 2011, high-quality coffee accounted for about 75%, while in 2012, low-quality coffee exceeded 60%. This indicates that the harvest year has little to do with coffee quality and has limited impact on predicting quality.

Exploratory analysis focuses on multidimensional visualization of coffee quality related variables, clarifying the direction of variable selection for subsequent modeling.

# FDA Part

# Initial logistic regression model

```{r}
#| echo: false
#| warning: false
#| message: false
#| tbl-cap: coefficient of full model
model_full <- glm(
  Qualityclass ~ aroma + flavor + acidity +
                category_two_defects +
                altitude_km +
                harvested,
  family = binomial(link = "logit"),
  data = coffeedata
)

library(broom)

coef_table <- tidy(model_full) %>%
  select(term, estimate, std.error, statistic, p.value) %>%
  rename(
    Term = term,
    Estimate = estimate,
    `Std. Error` = std.error,
    `z value` = statistic,
    `p value` = p.value
  )

coef_table %>%
  kable(
    caption = "Coefficient estimates for the initial logistic regression model",
    align = "ccccc",
    booktabs = TRUE,
    digits = 3
  ) %>%
  kable_styling(
    latex_options = "hold_position",
    font_size = 11
  ) %>%
  row_spec(0, align = "center")

```

A binomial GLM with logit link was fitted to investigate how sensory attributes, defect counts, altitude groups and harvest year influence the probability of a coffee batch being classified as good quality.

To assess whether the predictors included in the initial logistic regression model exhibit problematic multicollinearity, Variance Inflation Factors (VIFs) were calculated. This is particularly relevant because the exploratory analysis suggested relatively strong positive correlations among aroma, flavor and acidity.

# Model selection

```{r}
# stepwise selecting model via AIC
null_model <- glm(Qualityclass ~ 1,
                  data = coffeedata,
                  family = binomial)
step_model <- step(null_model, 
                   scope = list(lower = null_model, upper = model_full),
                   direction = "both",
                   trace = TRUE)
summary(step_model)
```

# Multicollinearity check

```{r}
#| echo: false
#| warning: false
#| message: false
library(car)
vif_values <- vif(step_model)
vif_table <- data.frame(
  Variable = names(vif_values),
  VIF=as.numeric(vif_values)
)

vif_table %>%
  kable(
    caption = "Variance Inflation Factors for the initial logistic regression model",
    align = "cccc",
    booktabs = TRUE,
    digits = 3
  ) %>%
  kable_styling(
    latex_options = "hold_position",
    font_size = 11
  ) %>%
  row_spec(0, align = "center")
```

Table X presents the VIF values for all predictors included in the initial model. Overall, the sensory variables show comparatively larger VIF values than the remaining predictors, which is consistent with the correlation patterns observed in the exploratory analysis. These results suggest that multicollinearity should be taken into account in subsequent model refinement.

# Visualising model fit

## Fitted Probability vs Observed

```{r}
coffeedata$pred_prob <- predict(step_model, type = "response")

ggplot(coffeedata, aes(x = pred_prob, y = factor(Qualityclass))) +
  geom_jitter(height = 0.1, alpha = 0.4) +
  labs(
    title = "Predicted probability vs observed coffee quality",
    x = "Predicted probability of good coffee",
    y = "Observed quality class"
  ) +
  theme_minimal()
```

## Residual Plot

```{r}
res <- residuals(step_model, type = "deviance")
fit <- fitted(step_model)

ggplot(data.frame(fit, res), aes(x = fit, y = res)) +
  geom_point(alpha = 0.4) +
  geom_smooth(se = FALSE) +
  labs(
    title = "Deviance residuals vs fitted values",
    x = "Fitted probability",
    y = "Deviance residuals"
  ) +
  theme_minimal()
```

# Interpretation of model coefficients

## Odds ratio

```{r}
#| echo: false
#| warning: false
#| message: false
library(sjPlot)

# Coefficient plot with CIs
plot_model(step_model, type = "est", show.values = TRUE)
```
